---
title: "Exploring Linguistic Biases"
date: "`r Sys.Date()`"
bibliography: references.bib
biblio-style: apalike
---

## Project title

Exploring Linguistic Biases 

### Research area and problem
The research area I am interested in learning more about is sociolinguistics, which is the study of language in society and how it interacts with characteristics like gender, perceptions, attitudes, and geographic region. A current topic of debate that I would like to address is where linguistic attitudes and biases come from in the United States. Linguistic discrimination affects many individuals in the United States because they are treated differently based on how they speak. There is still information to be learned about the patterns and origins of linguistic discrimination, in hopes of figuring out what we can do to prevent it.

### Research aim and question
The aim of this research will be to use exploratory data analysis to investigate the origins and examples of linguistic discrimination on social media in the United States. I will investigate and analyze discriminatory language on Twitter in hopes of understanding where certain biases are more prevalent and what contexts they are linked to. The research questions I would like to address are where are certain linguistic biases most prevalent, and in which contexts do they occur? 

## Reproduce


### Session

<details><summary>View session information</summary>

```{r}
sessioninfo::session_info()
```

</details>

```{r cleanup, echo=FALSE}
rm(list = ls()) # clean working environment
```

## References
